Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2015-07-12 Reviews LMCS/Response.txt

We thank the referees for their careful reading of our paper, and for the detailed and helpful feedback they have provided. Here, we will discuss each review in turn, and highlight the changes we have made in response. For the suggested minor changes, we have implemented most, and will comment only on those we chose not to.

@@

p.25, Definition 37: I find this definition a bit disturbing, it seems ad hoc at first and leaves one wondering whether something is wrong with your encoding. [...]

*** We agree that Definition 37 is not well-explained. We have opted for the following solution: we now explicitly define the translation of an instance of NCL-reconfiguration (which consists of a constraint graph with two configurations).

p.30, Theorem 46: just a minor quibble: you could make explicit the reduction from NCG-reconfiguration to the proof net equivalence problem, in order for the reader to better understand the complexity of such a reduction, which must be at least polytime computable. [...]

*** We have also made this translation more explicit (in the final theorem as well as earlier in the paper).

@@

Review 2

I have the impression that the authors do not realize the importance of their own result. [...]

*** While we were aware that MELL (without units) may encode MLL, we indeed did not realise that this consequence of our result (that MELL has no canonical normal forms) would be unexpected, or considered more interesting than the consequence for MLL. And we remain slightly cautious: surely proof equivalence for MELL is completely dominated by the complexity of cut-elimination? We have added a paragraph to this effect to the introduction.

### 2015-07-12 main/LMCS.tex
Remove dubious statement
I don’t think this really makes sense as stated, and I can’t think
of a nice way to phrase it correctly. I think it’s fine without.
more edits; final author response

\begin{abstract}
\capsabbrev{mll} proof equivalence is the problem of deciding whether two proofs in multiplicative linear logic are related by a series of inference permutations.
%
It is also known as the word problem for $*$-autonomous categories.
%
Previous work has shown the problem to be equivalent to a rewiring problem on proof nets, which are not canonical for full \capsabbrev{mll} due to the presence of the two units.
%
Drawing from recent work on reconfiguration problems, in this paper it is shown that \capsabbrev{mll} proof equivalence is \capsabbrev{pspace}-complete, using a reduction from Nondeterministic Constraint Logic.
%
An important consequence of the result is that the existence of a satisfactory notion of proof nets for \capsabbrev{mll} with units is ruled out (under current complexity assumptions).
%
Another is that for \capsabbrev{mell}, even without units, equivalence of normal forms is \capsabbrev{pspace}-hard.
\end{abstract}

@@

Here, we will introduce a similar necessary condition for the equivalence of two proof nets. Consider the example of the identity and twist maps on $\bot\tn\bot$, below. The two maps are semantically distinct; hence they are often given as a minimal example showing the necessity of having jumps in proof nets.
\[
	\twistid0 \qquad\qquad \twistid1
\]
In this case, it is easily seen that neither jump can be rewired, and that the two nets are thus inequivalent; to change from one net to the other, two jumps would need to be rewired simultaneously. This prompts the question of what would happen in a larger net: would a complicated series of rewirings be able to exchange two jumps configured like those of the above nets? Or phrased semantically: are there maps that equalise the identity and twist on $\bot\tn\bot$ in the free $*$-autonomous category?

@@

\capsabbrev{pspace}-hardness is by the encoding of \capsabbrev{ncg}-reconfiguration. Recall that an instance of \capsabbrev{ncg}-reconfiguration $(G,\gamma,\delta)$ is encoded by the triple $(\itn G,\itn\gamma,\itn\delta_\gamma)$. Then:
\[
	\gamma\perm\delta \qquad\iff\qquad \itn\gamma\perm\itn\delta_\gamma~.
\]
%
The direction $(\Rightarrow)$ is by Lemma~\ref{lem:completeness}; the direction $(\Leftarrow)$ is by Lemma~\ref{lem:soundness} and Proposition \ref{prop:retraction}.
\end{proof}

\subsection*{Acknowledgements}
We would like to thank Beniamino Accattoli, Dominic Hughes, and Lutz Stra{\ss}burger for helpful discussions and comments. We would also like to thank the anonymous referees for insightful remarks.

>>>>

\begin{abstract}
\capsabbrev{mll} proof equivalence is the problem of deciding whether two proofs in multiplicative linear logic are related by a series of inference permutations.
%
It is also known as the word problem for $*$-autonomous categories.
%
Previous work has shown the problem to be equivalent to a rewiring problem on proof nets, which are not canonical for full \capsabbrev{mll} due to the presence of the two units.
%
Drawing from recent work on reconfiguration problems, in this paper it is shown that \capsabbrev{mll} proof equivalence is \capsabbrev{pspace}-complete, using a reduction from Nondeterministic Constraint Logic.
%
An important consequence of the result is that the existence of a satisfactory notion of proof nets for \capsabbrev{mll} with units is ruled out (under current complexity assumptions).
%
The \capsabbrev{pspace}-hardness result extends to equivalence of normal forms in \capsabbrev{mell} without units, where the weakening rule for the ``why-not'' exponential induces a similar rewiring problem.
\end{abstract}

@@

Here, we will introduce a similar necessary condition for the equivalence of two proof nets. Consider the example of the identity and twist maps on $\bot\tn\bot$, below. The two maps are semantically distinct; hence they are often given as a minimal example showing the necessity of having jumps in proof nets.
\[
	\twistid0 \qquad\qquad \twistid1
\]
In this case, it is easily seen that neither jump can be rewired, and that the two nets are thus inequivalent; to change from one net to the other, two jumps would need to be rewired simultaneously. This prompts the question of what would happen in a larger net: would a complicated series of rewirings be able to exchange two jumps configured like those of the above nets?

@@

\capsabbrev{pspace}-hardness is by the encoding of \capsabbrev{ncg}-reconfiguration. Recall that an instance of \capsabbrev{ncg}-reconfiguration $(G,\gamma,\delta)$ is encoded by the triple $(\itn G,\itn\gamma,\itn\delta_\gamma)$. Then:
\[
	\gamma\perm\delta \qquad\iff\qquad \itn\gamma\perm\itn\delta_\gamma~.
\]
%
The direction $(\Rightarrow)$ is by Lemma~\ref{lem:completeness}; the direction $(\Leftarrow)$ is by Lemma~\ref{lem:soundness} and Proposition \ref{prop:retraction}. The encoding involves computing $\itn G$, $\itn\gamma$, and $\itn\delta$, then computing the parity of $\itn\gamma$ and $\itn\delta$ and adjusting $\itn\delta$ if needed; each can be done in polynomial time in the size of $(G,\gamma,\delta)$.
\end{proof}

\subsection*{Acknowledgements}
We would like to thank Beniamino Accattoli, Dominic Hughes, and Lutz Stra{\ss}burger for helpful discussions and comments. We would also like to thank the anonymous referees for their insightful remarks.
